---
title: "metabolism"
author: "mcdevitt"
date: "26 mars 2017"
output:
  html_document:
    highlight: tango
    theme: united
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Metabolism - Analysis

The Metabolic data on Canvas has the average mass, metabolic rate and average lifespan of 95 different species of mammals.  Kleiber’s law states that the metabolic rate of an animal species, on average, is proportional to its mass raised to the power of ¾.  Judge the adequacy of this theory with these data.  

```{r metabolism, echo = FALSE}


# ...	read in data

home_dir <- "~/_smu/_src/metabolism/"
setwd(home_dir)
data_dir <- "./data"

setwd(data_dir)

s26 <- read.csv("ex0826.csv", stringsAsFactors = FALSE)

setwd(home_dir)

# ...	add some additional columns for log transforms
exp <- 0.75
s26$m34 <- s26$Mass^exp
s26$log_mass_75 <- log(s26$m34)
s26$log_metab <- log(s26$Metab)

```


### __Strucutre of Data Set (with added columns for log transforms)__  

```{r}

str(s26)

```

#### __Add some basic plots for data exploration__  
```{r}
plot(s26$Metab ~ s26$Mass,
	 main = "Metabolic Rates vs. Average Mass of 95 mammal species",
	 xlab = "Mass",
	 ylab = "Metabolic Rate")
```

##### Data set shows very right skewed distribution in both x and y axes as evidenced by the data points in the upper right corner of the plot. In order to reduce the effect of the skewness in the raw data set, the data is transformed to log scale on both x & y axes.  

```{r}
plot(s26$Metab ~ s26$Mass, log = "xy",
	 main = "Metabolic Rates vs. Average Mass of 95 mammal species",
	 xlab = "log(Mass)",
	 ylab = "log(Metabolic Rate)")
```

##### Since the base question is to evaluate if the data set is consistent with Kleiber's law, an additional transformation is included to raise the Mass to the 3/4 power. This data plot shows similar apparently normal distributions in both x and y axes, and sufficiently linear to proceed with the regression analysis with the data transformed in this manner.

```{r}
plot(s26$Metab ~ s26$m34, log = "xy",
	 main = "Metabolic Rates vs. Average Mass of 95 mammal species",
	 xlab = "log(Mass^3/4)",
	 ylab = "log(Metabolic Rate)")

# ... define data set to analyze 

df_xy <- data.frame(s26$log_mass_75, s26$log_metab)

names(df_xy) <- c("log_mass_75", "log_metab")

str(df_xy)

```

### Homework 11 - Problem Set 1 - Metabolism

#### b. Analyze the data providing at least the following:    

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__i. A Scatterplot with confidence intervals of the regression line and prediction intervals of the regression line.  Please do in SAS and R!__    
```{r, echo = FALSE}


fit_all <- lm(log_metab ~ log_mass_75, data = df_xy)

df_xy_conf_int <- as.data.frame(predict(fit_all, df_xy , interval = "confidence"))

plot(df_xy$log_metab ~ df_xy$log_mass_75, col = 'blue', pch = 19)
polygon(c(df_xy$log_mass_75, rev(df_xy$log_mass_75)),
		c(df_xy_conf_int$lwr, rev(df_xy_conf_int$upr)),col="lightgrey")

lines(fit_all$fitted.values ~ df_xy$log_mass_75, col = 'black', lwd = 4, lty = 3)

lines(df_xy_conf_int$lwr ~ df_xy$log_mass_75, col = 'black', lwd = 1)
lines(df_xy_conf_int$upr ~ df_xy$log_mass_75, col = 'black', lwd = 1)

points(df_xy$log_metab ~ df_xy$log_mass_75, col = 'blue', pch = 19)

df_xy_pred_int <- as.data.frame(predict(fit_all, df_xy, interval = "predict"))

lines(df_xy_pred_int$lwr ~ df_xy$log_mass_75, col = 'grey', lwd = 1, lty = 2)
lines(df_xy_pred_int$upr ~ df_xy$log_mass_75, col = 'grey', lwd = 1, lty = 2)

```

And the SAS plot ...

  
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__ii. A table showing the t-statistics and pvalues for the significance of the regression parameters: .   Please do in SAS and R!__  

```{r plot, echo=FALSE}

summary(fit_all)

```

And the SAS table ...

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__iii. Using the data in ii show all 6 steps of each hypothesis test.__   
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  

##### __Six-Step Hypothesis Test - Slope__   

``` {r, echo = FALSE}

alpha <- 0.05
dof <- dim(df_xy)[1] - 2
crit_value <- qt(1 - alpha/2, dim(df_xy)[1] - 2)

t_value <- coef(summary(fit_all))["log_mass_75","t value"]
p_value <- coef(summary(fit_all))["log_mass_75","Pr(>|t|)"]

if (p_value < alpha)
{
	acc_rej <- "Reject Ho"
	is_is_not <- "is"
	does_does_not <- "does not"
} else {
	acc_rej <- "Do not Reject Ho"
	is_is_not <- "is not"
	does_does_not <- "does"
}

cnf_intrvl <- confint(fit_all, 'log_mass_75', level = 1 - alpha)
cnf_intrvl_lwr <- cnf_intrvl[1,1]
cnf_intrvl_upr <- cnf_intrvl[1,2]

print_cnf_intrvl <- paste0(sprintf("(%9.3f", cnf_intrvl_lwr), sprintf(", %9.3f)", cnf_intrvl_upr))

```

1: Ho : b1 = 0; Ha ; b1 != 0  
2 : Critical Value : t(0.975, df = __`r dof `__) = +/- __`r crit_value`__   
3 : t = `r t_value `    
4 : p_value = `r p_value` < 0.05 ?    
5: `r acc_rej`  
6 : There `r is_is_not` sufficient evidence to suggest at the alpha = `r alpha` level of significance (p-value = `r p_value`) that the slope of the regression line that estimates the T cell for Metabolism based on the log_mass_75 of the birds `r is_is_not` equal to zero. A `r (1 - alpha)*100`% confidence interval for the slope is `r print_cnf_intrvl`, which is an interval that `r does_does_not` contain the value zero  
  

##### __Six-Step Hypothesis Test - Intercept__   

``` {r, echo = FALSE}

alpha <- 0.05
dof <- dim(df_xy)[1] - 2
crit_value <- qt(1 - alpha/2, dim(df_xy)[1] - 2)

t_value <- coef(summary(fit_all))["(Intercept)","t value"]
p_value <- coef(summary(fit_all))["(Intercept)","Pr(>|t|)"]

if (p_value < alpha)
{
	acc_rej <- "Reject Ho"
	is_is_not <- "is"
	does_does_not <- "does not"
} else {
	acc_rej <- "Do not Reject Ho"
	is_is_not <- "is not"
	does_does_not <- "does"
}

cnf_intrvl <- confint(fit_all, '(Intercept)', level = 1 - alpha)
cnf_intrvl_lwr <- cnf_intrvl[1,1]
cnf_intrvl_upr <- cnf_intrvl[1,2]

print_cnf_intrvl <- paste0(sprintf("(%9.3f", cnf_intrvl_lwr), sprintf(", %9.3f)", cnf_intrvl_upr))

```

1 : Ho : b0 = 0; Ha ; b0 != 0  
2 : Critical Value : t(0.975, df = __`r dof `__) = +/- __`r crit_value`__   
3 : t = `r t_value `    
4 : p_value = `r p_value` < 0.05 ?    
5: `r acc_rej`  
6 : There `r is_is_not` sufficient evidence to suggest at the alpha = `r alpha` level of significance (p-value = `r p_value`) that the intercept of the regression line that estimates the T cell for Metabolism based on the log_mass_75 of the birds is not equal to zero. A `r (1 - alpha)*100`% confidence interval for the slope is `r print_cnf_intrvl`, which is an interval that `r does_does_not` contain the value zero  


=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__iv. The regression equation.__    


``` {r, echo = FALSE}

b0 <- fit_all$coefficients[1]
b1 <- fit_all$coefficients[2]

```

The Regression equation : 
* log_metab = `r b1` * log_mass_75 + `r b0`


=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__v. Interpretation of the slope and intercept in the model (regression equation.)__   

##### The slope represents the rate of change in log_metab estimated for a unit change in stone log_mass_75 transported by this populations of birds sampled. I.e., for each increase in 1 gram of stone log_mass_75 there is a corresponding increase in log_metab count of `r b1` 

##### The intercept represents the estimated average level of log_metab counts associated to a transported stone log_mass_75 of 0 grams by this population of birds. From a practical point of view, since each bird evalauted did actually transport stones of some log_mass_75, the intercept provides an estimate of the lower bound of log_metab counts in this population of birds.

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__vi. Find and interpret the 95% confidence interval for the mean t-cell response conditional on a stone log_mass_75 of 4.5 grams.  Please do in SAS and R!__  

``` {r, echo = FALSE} 

cnf_intrvl <- predict(fit_all, data.frame(log_mass_75 = 4.5) , interval = "confidence")
cnf_intrvl_lwr <- cnf_intrvl[1,2]
cnf_intrvl_upr <- cnf_intrvl[1,3]

print_cnf_intrvl <- paste0(sprintf("(%9.3f", cnf_intrvl_lwr), sprintf(", %9.3f)", cnf_intrvl_upr))

```


The confidence interval is the upper and lower bound for the expected mean value at the given indpendent value (e.g., log_mass_75 = 4.5 grams) for the current regression relationship.

For this particular regression, the confidence interval at stone log_mass_75 = 4.5 : log_metab count in the range `r print_cnf_intrvl`
  
=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__vii. Find and interpret the 95% prediction interval for the predicted t-cell response given a stone log_mass_75 of 4.5 grams.   Please do in SAS and R!__  


``` {r, echo = FALSE} 

prd_intrvl <- predict(fit_all, data.frame(log_mass_75 = 4.5) , interval = "predict")
prd_intrvl_lwr <- prd_intrvl[1,2]
prd_intrvl_upr <- prd_intrvl[1,3]

print_prd_intrvl <- paste0(sprintf("(%9.3f", prd_intrvl_lwr), sprintf(", %9.3f)", prd_intrvl_upr))

```


The prediction interval is the upper and lower bound for an 'next' observation of dependent value at the given independent value, based on the current regression relationshhip. That is to say, for an observation of log_mass_75 and log_metab not included in this analysis, the prediction interval bounds the range of future observations that are expected.  

For this particular case, the prediction interval at stone log_mass_75 = 4.5 : log_metab count in the range `r print_prd_intrvl`

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  

__And the SAS table__ ...  




=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__x. A scatterplot of residuals.    Please do in SAS and R!__  

``` {r, echo = FALSE}

log_metab_res = resid(fit_all)

plot(df_xy$log_mass_75, log_metab_res, 
     pch = 19,
	 col = "blue",
	 ylab="Residuals",
	 xlab="Stone log_mass_75", 
     main="Metabolism - Residual Plot") 
abline(0, 0)

```


=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=--=-=-=-  
__xi. A histogram of residuals with normal distribution superimposed.  (from SAS).__    


``` {r, echo = FALSE}

res_mean <- mean(log_metab_res)
res_std <- sqrt(var(log_metab_res))

hist(log_metab_res, density=20, breaks = 10, prob = TRUE, 
     xlab="Stone log_mass_75",
#	 ylim=c(0, 10), 
     main="Metabolism - Residuals")
curve(dnorm(x, mean = res_mean, sd = res_std),
	  col = "darkblue",
	  lwd = 2,
	  add = TRUE,
	  yaxt = "n")

```

And the SAS plots ...

